Clustering in a hyperbolic model of complex networks.

### 2019-02-05 concentration_argument.tex
updated notation and added comments to sections 1-3

Many of the computations in this paper will involve integrals whose integrand contains the function $\Prob{\Po(\mu_n(y)) = k}$, where $\Po(\lambda)$ denotes a Poisson random variable with expectation $\lambda$ and $\mu_n(y) \approx e^{y/2}$. For instance, the expected modified local clustering function $c_\infty(k)$ in $G_\Pcal$ equals
\[
	\frac{\int_0^\infty \Prob{\Po(\Mu{\BallPo{0,y}}) = k} \Delta_{\Pcal}(y) \alpha e^{-\alpha y} \dd y}
	{\int_0^\infty \Prob{\Po(\Mu{\BallPo{0,y}}) = k} \alpha e^{-\alpha y} \dd y},
\]
where $\Mu{\BallPo{0,y}} = \xi_{\alpha,\nu} e^{y/2}$ and $\Delta_{\Pcal}(y)$ is defined by~\eqref{eq:def_delta_p}.

In particular, we already established that the degree of a node at height $y$ in the infinite model is Poisson with mean $\xi_{\alpha,\nu} e^{y/2}$. Since we are mostly interested in those nodes with degree $k_n$ and Poisson random variables are well concentrated around their mean, we would therefore like to be able the restrict integration to intervals where $e^{y/2} \approx k_n$. Moreover, we want to be able to do this for all three Poisson models $G_{\HP,n}$, $G_{\Pcal,n}$ and $G_\Pcal$ as well as simultaneously deal with the case where $k_n \to \infty$ and $k_n = \bigT{1}$. In this section we will establish such a result. We start with a concentration lemma for the infinite model (Lemma~\ref{lem:concentration_argument}) and explain in Remark~\ref{rmk:concentration_argument} how such a result will be used throughout the paper. To obtain similar results for the other two models we have to first analyze the average degree in these models which we do in Sections~\ref{ssec:average_degree_HP_n} and~\ref{ssec:average_degree_P_n}. We conclude this section with a general result that allows us to extend the concentration lemma for the infinite model to the hyperbolic random graph and finite box model.

@@

Recall that $\rho(y, k)$ is the probability density function of a Poisson random variable with expectation $\mu_{\alpha, \nu}(B_\Pcal(y)) = \xi_{\alpha,\nu}e^{\frac{y}{2}}$. 
For any sequence $\{k_n\}_{n \ge 1}$, possibly constant, define
\begin{equation}\label{eq:def_kappa_n}
	\kappa_n := \begin{cases}
		\log(n) &\mbox{if } k_n = \bigT{1},\\
		\sqrt{k_n \log(k_n)} &\mbox{else.}
	\end{cases}
\end{equation}
\TM{ This is not a good definition. There are sequences that neither stay bounded nor tend to infinity, e.q.~odd values equal to 2 even values equal to $\log n$. How do you decide for a given $k_n$ which of the two cases you pick. To solve this we could either restrict to sequences that are either a) constant or b) tending to infinity; or we could make a cut-off at some specific, slow growing function such as $\log\log\log n$. }\PvdH{You are right, I overlooked this fact. I would opt to go with the setting where we consider only sequences that are either asymptotically bounded or tend to infinity. After the new proves are added we can see where we specify this.}
and define further, for any $C > 0$
\begin{equation}\label{eq:def_K_C_set}
	\Kcal_C(k_n) = \left\{p=(x,y) \in \R \times \R_+ : \frac{k_n - C \kappa_n}{\xi_{\alpha,\nu}} \le e^{\frac{y}{2}}
	\le \frac{k_n + C \kappa_n}{\xi_{\alpha,\nu}} \right\},
\end{equation}
Where we recall that $a \vee b = \max\{a,b\}$ and $a \wedge b = \min\{a,b\}$.

Note that if $k_n = \Omega(\log(n))$ then $e^{y/2} = \bigT{k_n}$ whenever $p=(x,y) \in \Kcal_C(k_n)$. The next lemma states that for a large class of functions $h(y)$, to compute the integral 
\[
	\int_{\Rcal_n} \rho(y,k_n) h(y) f_{\alpha,\nu}(x,y) \dd x \dd y
\]
\TM{ This part is supposed to be for the infinite model, yet you integrate over the finite box $\Rcal_n$. Is this really meant?
I would have expected we just integrate wrt.~$y$ and replace $f_{\alpha,\nu}(x,y)$ by $\alpha e^{-\alpha y}$, the exponential density.}\PvdH{I understand this confusion and agree that your suggestion makes more sense here. However, I choose this setup to minimize notations later. I am for changing the setup to what you are suggesting but I will have to spend some time first on checking how that will effect the definitions in later sections.}
it is enough to consider integration \BL{over} $\Kcal_C(k_n)$ instead of $\Rcal_n$. More precisely, for any $\ell_n = (1 + \smallO{1})k_n$ it is enough to consider $\Kcal_C(\ell_n)$. The flexibility of using $(1+\smallO{1})k_n$ instead of $k_n$ will prove useful in later sections.

\begin{lemma}\label{lem:concentration_argument}
Let $\alpha > \frac{1}{2}$, $\nu > 0$, $\{k_n\}_{n \ge 1}$ be any positive sequence such that $k_n = o(n^{\frac{1}{2\alpha + 1}})$ and let $\ell_n = k_n(1 + \epsilon_n)$, with $\epsilon_n \to 0$. In addition let $\beta < \alpha$ and $h : \R_+ \rightarrow  \R$ be a any function such that $h(y) = g(y)e^{\beta y}$ with $|g(y)|$ uniformly bounded on $\R_+$. Then, if we define for $C > 0$
\[
	\lambda_n^\pm = (\ell_n \pm C \kappa_n) \wedge \kappa_n, 
	\quad a_n^\pm = 2 \log\left(\frac{\lambda_n^\pm}{\xi_{\alpha,\nu}}\right),
\] 
\TM{ If $k$ is constant, then $\ell^- < 0$ and hence 
$a^-$ is undefined! Need to change the definitions. }\PvdH{You are right. I changed it by taking the minimum with $\kappa_n$.}

we have
\begin{equation}\label{eq:error_bound_int_rho_not_K}
\begin{aligned}
	\int_{\Rcal_n\setminus \Kcal_{C}(k_n)} \rho(y,k_n) h(y) \alpha e^{-\alpha y} \dd x \dd y
	&= \begin{cases}
		\bigO{\log(n)^{k_n} n^{-\frac{C}{2}}} &\mbox{if } k_n = \bigT{1} \\
		\bigO{k_n^{-(1+C^2)/2}} &\mbox{else.}
	\end{cases}
\end{aligned}
\end{equation}
\TM{ As before, this is not a proper dichotomy. (unless we assume more about the sequences $k_n$, i.e.~either constant or tending to infinity }
as $n \to \infty$.

In particular, if $h_n(y)$ is a function such that $h_n(y) = \bigO{n^{-1} k_n^{s} e^{\beta y}}\rho(y,k_n)$ for some $s \in \R$ and $\beta < \alpha$ as $n \to \infty$. Then for $C > 0$ large enough we have
\[
	\lim_{n \to \infty} \int_{\Rcal_n \setminus \Kcal_{C}(\ell_n)} h_n(y) f_{\alpha,\nu}(x,y) \dd x\dd y = 0,
\]
or equivalently,
\[
	\int_{\Rcal_n} \hspace{-5pt} h_n(y) f_{\alpha,\nu}(x,y) \dd x \dd y
	= (1+\smallO{1}) \int_{\Kcal_{C}(\ell_n)} \hspace{-5pt} h_n(y) f_{\alpha,\nu}(x,y) \dd x \dd y.
\]
\end{lemma}

\begin{proof}
Note that in this case
\[
	\kappa_n := \begin{cases}
		\log(n) &\mbox{if } k_n = \bigT{1},\\
		\sqrt{\ell_n \log(\ell_n)} &\mbox{else.}
	\end{cases}
\]
Then, since $a_n^- < k_n$ and $a_n^+ > k_n$, $\rho_{y}(k_n)$, as a function of $y$, is strictly increasing on $[0,a_n^-]$ and strictly decreasing on $[a_n^+,\infty)$. 
\TM{ PLease give more details. }
Therefore, by our assumption on $h(y)$,
\begin{align*}
	&\hspace{-20pt}\int_{\R_+ \setminus [a_n^-, a_n^+]} h(y) \rho_{k_n}(y) 
		\alpha e^{-\alpha y} \dd y\\
    &= \bigO{1} \int_0^{a_n^-} e^{\beta y} \rho_{k_n}(y) \alpha e^{-\alpha y} \dd y 
    	+ \bigO{1}\int_{a_n^+}^{R_n} e^{\beta y} \rho_{k_n}(y) \alpha e^{-\alpha y}x \dd y \\
    &= \bigO{1} \int_0^{a_n^-} \rho_{k_n}(y) e^{-(\alpha-\beta) y} \dd y 
   		+ \bigO{1} \int_{a_n^+}^{R_n} \rho_{k_n}(y) e^{-(\alpha-\beta) y} \dd y\\
   	&\le \bigO{1}\rho(a_n^-,k_n)\int_0^{a_n^-} e^{-(\alpha-\beta) y} \, \dd y
   		+ \bigO{1} \rho(y,a_n^+) \int_{a_n^+}^{R_n} e^{-(\alpha-\beta) y} \, \dd y.
\end{align*}
Note that if $k_n = \bigT{1}$ then, for large enough $n$, the first integral is zero. We conclude that
\begin{equation}\label{eq:concentration_lemma_integral_bound}
\begin{aligned}
	\int_{\R_+ \setminus [a_n^-, a_n^+]} h(y) \rho_{k_n}(y) \alpha e^{-\alpha y} \dd y
	&= \begin{cases}
		\bigO{1} \rho(a_n^+,k_n) &\mbox{if } k_n = \bigT{1},\\
		\bigO{1} \left(\rho(a_n^-,k_n) + \rho(a_n^+,k_n)\right) &\mbox{else.}
	\end{cases}	
\end{aligned}
\end{equation}

We shall now bound the terms $\rho(a_n^\pm,k_n)$, starting with $\rho(a_n^+,k_n)$.  Using that $k! > \sqrt{2\pi} k^{k + 1/2} e^{-k}$ \TM{ Reference! } we write
\begin{align*}
	\rho(a_n^+,k_n) &\BL{=} \frac{\Mu{\BallPo{a_n^+}}^{k_n}}{k_n!} e^{-\Mu{\BallPo{a_n^+}}} \\
	&\le (2\pi)^{-1/2} k_n^{-1/2} \left(\frac{\Mu{\BallPo{a_n^+}}}{k_n}\right)^{k_n} e^{-(\Mu{\BallPo{a_n^+}} - k_n)}\\
	&= (2\pi)^{-1/2} k_n^{-1/2} 
		e^{-k_n\left(\frac{\Mu{\BallPo{a_n^+}}}{k_n} - 1 - \log\left(\frac{\Mu{\BallPo{a_n^+}}}{k_n}\right)\right)}.
\end{align*}

Let us first consider the case $k_n \to \infty$,
\TM{ again, in general we cannot neatly split into the two cases $k\to\infty$ and $k=\Theta(1)$ } in which case $\kappa_n = \sqrt{\ell_n \log(\ell_n)}$. Since 
\[
	\frac{\Mu{\BallPo{a_n^+}}}{k_n} = \frac{\lambda_n^{\BL{+}}}{k_n} = 1 + \epsilon_n + C \frac{\kappa_n}{k_n} 
	= 1 + \epsilon_n + C \sqrt{\frac{(1+\epsilon_n)\log((1+\epsilon_n)k_n)}{k_n}},
\]
and $x - \log(1 + x) \sim x^2/2$ as $x \to 0$, we get 
\begin{align*}
	\rho(a_n^+,k_n) 
	&\le \sqrt{2\pi} k_n^{-1/2} 
		e^{-k_n\left(\epsilon_n + C \frac{\kappa_n}{k_n} - \log\left(1 + \epsilon_n + C \frac{\kappa_n}{k_n}\right)\right)}\\
	&\sim (2\pi)^{-1/2} k_n^{-1/2} e^{-\frac{k_n \left(\epsilon_n + C \kappa_n/k_n\right)^2}{2}} \\
	&= \bigO{k_n^{-(1+C^2)/2}},		\numberthis \label{eq:concentration_lemma_bound_an+}
\end{align*}
where for the last line we used that
\[
	-k_n \frac{\left(\epsilon_n + C \kappa_n/k_n\right)^2}{2} = -\frac{C^2}{2}\log(k_n) + \bigT{1}.
\]
Let use now consider the case $k_n = \bigT{1}$. Then $\kappa_n = \log(n)$,
\[
	\frac{\Mu{\BallPo{a_n^+}}}{k_n} = \frac{\lambda_n^+}{k_n} = 1 + \epsilon_n + C \frac{\log(n)}{k_n},
\] 
and hence
\begin{equation}\label{eq:concentration_lemma_bound_an+_constant}
	\rho(a_n^-,k_n) \le \bigO{\log(n)^{k_n} n^{-C}}.
\end{equation}

Note that for $\rho(a_n^-,k_n)$ we only need to consider the case $k_n \to \infty$. A similar analysis as for $a_n^-$ yields
\begin{equation}\label{eq:concentration_lemma_bound_an-}
	\rho(a_n^-,k_n) \le \bigT{1} k_n^{-1/2} e^{-\frac{k_n \left(\epsilon_n - C \kappa_n/k_n\right)^2}{2}} = \bigO{k_n^{-(1+C^2)/2}}.
\end{equation}

Plugging~\eqref{eq:concentration_lemma_bound_an-}, \eqref{eq:concentration_lemma_bound_an+} and~\eqref{eq:concentration_lemma_bound_an+_constant} into~\eqref{eq:concentration_lemma_integral_bound} yields the result. The second statement immediately follows from the first by choosing a large enough $C$ and observing that
\[
	\int_{\Rcal_n \setminus \Kcal_{C}(\ell_n)} h_n(y) f_{\alpha,\nu}(x,y) \dd x\dd y
	\le n \int_{\R_+ \setminus [a_n^-, a_n^+]} h_n(y) \alpha e^{-\alpha y} \dd y.
\]
\end{proof}

\begin{remark}[Concentration argument]\label{rmk:concentration_argument}
\TM{ again, this is a loaded term. I would in the very least change to ``concentration of height'' everywhere }
Lemma~\ref{lem:concentration_argument} will prove very useful in the remainder of this paper since we often have to deal with integrands of the form $g_n(y) f_{\alpha,\nu}(x,y)$ where $g_n(y) = \bigO{n^{-1} k_n^{s}}e^{\beta y}\rho(y,k_n)$ for some $s \in \R$ and $\beta < \alpha$. In this case the lemma tells us that for a suitable $C > 0$ we only need to integrate over $\Kcal_C(k_n)$. In order words, we may always assume for $g_n(y)$ (for a penalty of $\smallO{1}$) that $k_n - C \kappa_n \le \xi_{\alpha,\nu} e^{y/2} \le k_n + C \kappa_n$. We will refer to this as a \emph{concentration argument}, e.g. \emph{by a concentration argument}
\[
	\int_{\Rcal_n} g_n(y) f_{\alpha, \nu}(x,y) \dd x \dd y 
	= (1+\smallO{1}) \int_{\Kcal_C(k_n)} g_n(y) f_{\alpha, \nu}(x,y) \dd x \dd y.
\]
\PvdH{ @All:It might be a good idea to add one more example of how the concentration argument is used in the paper. I someone finds a good example we use later on, please add it here.}
\end{remark}

>>>>

In the proof of Proposition~\ref{prop:asymp} we used a result that allowed us to restrict integration over $y$ to the interval $[a(k)^-, a(k)^+]$, with
\[
	a(k)^\pm = 2\log\left(\frac{1 \pm C \sqrt{k \log(k)}}{\xi} \vee 1\right).
\]
The reason for this was that the integrand included the function $\rho(y,k) = \Prob{\Po(\mu(y)) = k}$, where $\Po(\lambda)$ denotes a Poisson random variable with expectation $\lambda$ and $\mu(y) = \mu(B_\Pcal(y)) = \xi e^{y/2}$, and Poisson random variables are well concentrated around their mean, i.e. around those values of $y$ for which $\mu(y) \approx k$.

In the remainder of this paper we will often encounter integrands involving, for some $\mu_n(y)$, the function $\Prob{\Po(\mu_n(y)) = k}$. In these case we want to be able to restrict our integration around those heights $y$ for which $\mu_n(y) \approx k_n$. We will refer to as \emph{concentration of heights arguments}. In this section we establish such results. We start with a concentration of heights lemma for the infinite model $\Ginf$ (Lemma~\ref{lem:concentration_argument}) and explain in Remark~\ref{rmk:concentration_argument} how such a result will be used throughout the paper. To obtain similar results for the other two models we have to first analyze the expected number of nodes in a typical neighborhood in these models. We do in Sections~\ref{ssec:average_degree_HP_n} and~\ref{ssec:average_degree_P_n}. We conclude this section with a general result that allows us to extend the concentration lemma for the infinite model to the hyperbolic random graph and finite box model.

@@

%Recall that $\rho(y, k)$ is the probability density function of a Poisson random variable with expectation $\mu_{\alpha, \nu}(B_\Pcal(y)) = \xie^{\frac{y}{2}}$. 
%We will consider two different types of sequences $\{k_n\}_{n \ge 1}$, those that are asymptotically bounded and those that diverge. We define
%\begin{equation}\label{eq:def_kappa_n}
%	\kappa_n := \begin{cases}
%		\log(n) &\mbox{if } k_n = \bigT{1},\\
%		\sqrt{k_n \log(k_n)} &\mbox{if } k_n = \omega(1).
%	\end{cases}
%\end{equation}
%\TM{ This is not a good definition. There are sequences that neither stay bounded nor tend to infinity, e.q.~odd values equal to 2 even values equal to $\log n$. How do you decide for a given $k_n$ which of the two cases you pick. To solve this we could either restrict to sequences that are either a) constant or b) tending to infinity; or we could make a cut-off at some specific, slow growing function such as $\log\log\log n$. }\PvdH{You are right, I overlooked this fact. I would opt to go with the setting where we consider only sequences that are either asymptotically bounded or tend to infinity. After the new proves are added we can see where we specify this.}
For any positive sequence $a_n \to \infty$ and $C > 0$ we define
\begin{equation}\label{eq:def_K_C_set}
	\Kcal_C(a_n) = \left\{y \in \R_+ : \frac{a_n - C \sqrt{a_n \log(a_n)}}{\xi} \vee 1 \le e^{\frac{y}{2}}
	\le \frac{a_n + C \sqrt{a_n \log(a_n)}}{\xi} \right\}.
\end{equation}
In addition we define
\begin{equation}\label{eq:def_K_C_n_set}
	\Kcal_{C,n}(a_n) := (-I_n, I_n] \times ((0,R_n] \cap \Kcal_{C}(a_n)),
\end{equation}
where $I_n := \frac{\pi}{2}e^{R_n/2}$. In addition, for a sequence $a_n \to \infty$, we shall often use the shorthand notation
\begin{equation}\label{eq:def_kappa_n}
	\kappa_n := \sqrt{a_n \log(a_n)}
\end{equation}

%Where we recall that $a \vee b = \max\{a,b\}$ and $a \wedge b = \min\{a,b\}$.

The next lemma states that for a large class of functions $h(y)$, to compute the integral 
\[
	\int_{0}^\infty \rho(y,k_n) h(y) e^{-\alpha y} \dd y
\]
%\TM{ This part is supposed to be for the infinite model, yet you integrate over the finite box $\Rcal_n$. Is this really meant?
%I would have expected we just integrate wrt.~$y$ and replace $f_{\alpha,\nu}(x,y)$ by $\alpha e^{-\alpha y}$, the exponential density.}\PvdH{I understand this confusion and agree that your suggestion makes more sense here. However, I choose this setup to minimize notations later. I am for changing the setup to what you are suggesting but I will have to spend some time first on checking how that will effect the definitions in later sections.}
it is enough to consider integration over $\Kcal_C(k_n)$ instead of $\R_+$. More precisely, for any $\ell_n = (1 + \smallO{1})k_n$ it is enough to consider $\Kcal_C(\ell_n)$. The flexibility of using $(1+\smallO{1})k_n$ instead of $k_n$ will prove useful in later sections.

\begin{lemma}\label{lem:concentration_argument}
Let $\alpha > \frac{1}{2}$, $\nu > 0$, $\{k_n\}_{n \ge 1}$ be any positive sequence such that $k_n = o(n^{\frac{1}{2\alpha + 1}})$ and let $\ell_n = k_n(1 + \epsilon_n)$, with $\epsilon_n \to 0$. In addition let $\beta < \alpha$ and $h : \R_+ \rightarrow  \R$ be a any function such that $h(y) = g(y)e^{\beta y}$ with $|g(y)|$ uniformly bounded on $\R_+$. Then, if we define for $C > 0$
\[
	\lambda_n^\pm = (\ell_n \pm C \sqrt{\ell_n \log(\ell_n)}) \wedge \xi, 
	\quad a_n^\pm = 2 \log\left(\frac{\lambda_n^\pm}{\xi}\right),
\] 
%\TM{ If $k$ is constant, then $\ell^- < 0$ and hence 
%$a^-$ is undefined! Need to change the definitions. }\PvdH{You are right. I changed it by taking the minimum with $\xi$.}
we have
\begin{equation}\label{eq:error_bound_int_rho_not_K}
	\int_{\R_+ \setminus \Kcal_{C}(\ell_n)} \rho(y,k_n) h(y) \alpha e^{-\alpha y} \dd y
	= \bigO{k_n^{-(1+C^2)/2}}
\end{equation}
%\TM{ As before, this is not a proper dichotomy. (unless we assume more about the sequences $k_n$, i.e.~either constant or tending to infinity }
as $n \to \infty$.

In particular, if $h_n(y)$ is a function such that $h_n(y) = \bigO{n^{-1} k_n^{s} e^{\beta y}}\rho(y,k_n)$ for some $s \in \R$ and $\beta < \alpha$ as $n \to \infty$. Then for $C > 0$ large enough we have
\[
	\lim_{n \to \infty} \int_{\Rcal_n \setminus \Kcal_{C,n}(\ell_n)} h_n(y) 
	f_{\alpha, \nu}(x,y) \dd x \dd y = 0,
\]
or equivalently,
\[
	\int_{\Rcal_n} \hspace{-5pt} h_n(y) f_{\alpha, \nu}(x,y) \dd x \dd y
	= (1+\smallO{1}) \int_{\Kcal_{C,n}(\ell_n)} \hspace{-5pt} h_n(y) f_{\alpha, \nu}(x,y) \dd x \dd y.
\]
\end{lemma}

\begin{proof}

Note that in this case we write
\[
	\kappa_n = \sqrt{\ell_n \log(\ell_n)}.
\]
and recall (see proof of Proposition~\ref{prop:asymp}) that $\rho(y,k_n)$, as a function of $y$, is strictly increasing on $[0,a_n^-]$ and strictly decreasing on $[a_n^+,\infty)$. Therefore, by our assumption on $h(y)$,
\begin{align*}
	&\hspace{-20pt}\int_{\R_+ \setminus [a_n^-, a_n^+]} h(y) \rho(y,k_n) 
		\alpha e^{-\alpha y} \dd y\\
    &= \bigO{1} \int_0^{a_n^-} e^{\beta y} \rho(y,k_n) \alpha e^{-\alpha y} \dd y 
    	+ \bigO{1}\int_{a_n^+}^{\infty} e^{\beta y} \rho(y,k_n) \alpha e^{-\alpha y}x \dd y \\
    &= \bigO{1} \int_0^{a_n^-} \rho(y,k_n) e^{-(\alpha-\beta) y} \dd y 
   		+ \bigO{1} \int_{a_n^+}^{\infty} \rho(y,k_n) e^{-(\alpha-\beta) y} \dd y\\
   	&\le \bigO{1}\rho(a_n^-,k_n)\int_0^{a_n^-} e^{-(\alpha-\beta) y} \, \dd y
   		+ \bigO{1} \rho(a_n^+,k_n) \int_{a_n^+}^{\infty} e^{-(\alpha-\beta) y} \, \dd y.
\end{align*}
Since $\alpha - \beta > 0$, we conclude that
\begin{equation}\label{eq:concentration_lemma_integral_bound}
	\int_{\R_+ \setminus [a_n^-, a_n^+]} h(y) \rho(y,k_n) \alpha e^{-\alpha y} \dd y
	= \bigO{1} \left(\rho(a_n^-,k_n) + \rho(a_n^+,k_n)\right) 
\end{equation}

We shall now bound the terms $\rho(a_n^\pm,k_n)$, starting with $\rho(a_n^+,k_n)$.  Using Stirling's approximation $k! \sim \sqrt{2\pi} k^{k + 1/2} e^{-k}$ as $k \to \infty$ we write
\begin{align*}
	\rho(a_n^+,k_n) &= \frac{\mu(a_n^+)^{k_n}}{k_n!} e^{-\mu(a_n^+)} \\
	&\sim (2\pi)^{-1/2} k_n^{-1/2} \left(\frac{\mu(a_n^+)}{k_n}\right)^{k_n} e^{-(\mu(a_n^+) - k_n)}\\
	&= (2\pi)^{-1/2} k_n^{-1/2} 
		e^{-k_n\left(\frac{\mu(a_n^+)}{k_n} - 1 - \log\left(\frac{\mu(a_n^+)}{k_n}\right)\right)}.
\end{align*}

Since 
\[
	\frac{\mu(a_n^+)}{k_n} = \frac{\lambda_n^{+}}{k_n} = 1 + \epsilon_n + C \frac{\kappa_n}{k_n} 
	= 1 + \epsilon_n + C \sqrt{\frac{(1+\epsilon_n)\log((1+\epsilon_n)k_n)}{k_n}},
\]
and $x - \log(1 + x) \sim x^2/2$ as $x \to 0$, we get 
\begin{align*}
	\rho(a_n^+,k_n) 
	&\le \sqrt{2\pi} k_n^{-1/2} 
		e^{-k_n\left(\epsilon_n + C \frac{\kappa_n}{k_n} - \log\left(1 + \epsilon_n + C \frac{\kappa_n}{k_n}\right)\right)}\\
	&\sim (2\pi)^{-1/2} k_n^{-1/2} e^{-\frac{k_n \left(\epsilon_n + C \kappa_n/k_n\right)^2}{2}} \\
	&= \bigO{k_n^{-(1+C^2)/2}},		\numberthis \label{eq:concentration_lemma_bound_an+}
\end{align*}
where for the last line we used that
\[
	-k_n \frac{\left(\epsilon_n + C \kappa_n/k_n\right)^2}{2} = -\frac{C^2}{2}\log(k_n) + \bigT{1}.
\]
A similar analysis as above yields
\begin{equation}\label{eq:concentration_lemma_bound_an-}
	\rho(a_n^-,k_n) \le \bigT{1} k_n^{-1/2} e^{-\frac{k_n \left(\epsilon_n - C \kappa_n/k_n\right)^2}{2}} = \bigO{k_n^{-(1+C^2)/2}}.
\end{equation}

Plugging~\eqref{eq:concentration_lemma_bound_an-} and~\eqref{eq:concentration_lemma_bound_an+}  into~\eqref{eq:concentration_lemma_integral_bound} yields the result. The second statement immediately follows from the first by choosing a large enough $C$ and observing that
\[
	\int_{\Rcal_n \setminus \Kcal_{C,n}(\ell_n)} h_n(y) f_{\alpha,\nu}(x,y) \dd x \dd y
	\le n \int_{\R_+ \setminus [a_n^-, a_n^+]} h_n(y) \alpha e^{-\alpha y} \dd y.
\]
\end{proof}

\begin{remark}[Concentration of heights argument]\label{rmk:concentration_argument}
%\TM{ again, this is a loaded term. I would in the very least change to ``concentration of height'' everywhere }
Lemma~\ref{lem:concentration_argument} will prove very useful in the remainder of this paper since we often have to deal with integrands of the form $g_n(y) f_{\alpha,\nu}(x,y)$ where $g_n(y) = \bigO{n^{-1} k_n^{s}}e^{\beta y}\rho(y,k_n)$ for some $s \in \R$ and $\beta < \alpha$. In this case the lemma tells us that for a suitable $C > 0$ we only need to integrate over $\Kcal_C(k_n)$. In order words, we may always assume for $g_n(y)$ (for a penalty of $\smallO{1}$) that $k_n - C \kappa_n \le \xi e^{y/2} \le k_n + C \kappa_n$. We will refer to this as a \emph{concentration of heights argument}, e.g. \emph{by a concentration of heights argument}
\[
	\int_{\Rcal_n} g_n(y) \rho(y,k_n) f_{\alpha,\nu}(x,y) \dd x \dd y 
	= (1+\smallO{1}) n \int_{\Kcal_C(k_n)} g_n(y) \rho(y,k_n) \alpha e^{-\alpha y} \dd y.
\]
\PvdH{ @All:It might be a good idea to add one more example of how the concentration argument is used in the paper. I someone finds a good example we use later on, please add it here.}
\end{remark}

We will later establish similar results for the cases where we consider the hyperbolic and finite box model. Here we have to consider Poisson distributions with mean $\mu(\BallHyp{y})$ and $\mu(\BallPon{y})$, respectively. For this the following, slightly more general version of Lemma~\ref{lem:concentration_argument} will be important.

\begin{lemma}\label{lem:concentration_argument_rho_approximation}
Let $\alpha > \frac{1}{2}, \nu > 0$, $k_n$ be any positive sequence such that $k_n = o(n^{\frac{1}{2\alpha + 1}})$, $\ell_n = (1 + \epsilon_n)k_n$, with $\epsilon_n \to 0$ and let $\Kcal_{C,n}(\ell_n)$ be defined as in~\eqref{eq:def_K_C_n_set}. In addition, define $\hat{\rho}_n(y,k) = \Prob{\Po(\hat{\mu}_n(y)) = k}$, where $\hat{\mu}_n(y)$ satisfies, 
\[
	\hat{\mu}_n(y) = (1 + \phi_n(y))\mu(y),
\]
where $\phi_n(y)$ is a continuous differentiable function such that for some $0 < \varepsilon < 1$
\[
	\sup_{0 \le y \le (1 - \varepsilon)R_n} |\phi_n(y)| = 0.
\]
Then the following holds for some $C > 0$ large enough:
\begin{enumerate}
\item For any sequence of continuous functions $h_n(y) : \R_+ \to \R$ such that, uniformly on $(0,R_n]$, $h_n(y) = \bigO{k_n^s e^{\beta y} \hat{\rho}_n(y,k_n)}$, as $n \to \infty$, 
%\TM{ Why not adjust $h$ so that we incorporate $\hat\rho$ in line below. Then both items look more similar. }\PvdH{The reason for the bound on $h_n(y)$ in the first statement is that we want to use this for cases where $h_n(y) \le k_n^s \hat{\rho}_n(y,k_n)$ (see for instance the proof of Proposition~\ref{prop:convergence_average_clustering_P_n}.}  
for some $s < 2\alpha(2\alpha + 1)$ and $\beta < \alpha$, we have,
\[
	\int_{\Rcal_n} h_n(y) f_{\alpha,\nu}(x,y) \dd x \dd y 
	= (1 + o(1)) \int_{\Kcal_{C,n}(\ell_n)} h_n(y) f_{\alpha,\nu}(x,y) \dd x \dd y.
\]
\item For any continuous bounded function $h(y)$,
\[
	\int_{\Kcal_{C,n}(\ell_n)} h(y) \hat{\rho}_n(y,k_n) f_{\alpha,\nu}(x,y) \dd x \dd y
	= (1+\smallO{1}) n \int_{0}^\infty h(y) \rho(y,k_n) \alpha e^{-\alpha y} \dd y,
\]
as $n \to \infty$.
\end{enumerate}
\end{lemma}

\begin{proof}
Similarly to the proof of Lemma~\ref{lem:concentration_argument} we define 
\[
	\lambda_n^\pm = (\ell_n \pm C \kappa_n) \wedge \xi, \quad \text{and} \quad a_n^\pm = 2 \log\left(\frac{\lambda_n^\pm}{\xi}\right).
\] 
%\TM{ Same problem as in L3.1. $\lambda^-$ can be negative, in which case $a^-$ is undefined.}

In addition we fix $\varepsilon^\ast > 0$ such that $\varepsilon < \min\{1- \frac{s}{2\alpha(2\alpha + 1)},\varepsilon, 1\}$ and. Then since $h_n(y) = \hat{\rho}_n(y,k_n) \bigO{k_n^s}$ we have
\begin{align*}
	k_n^s \int_{(1-\varepsilon^\ast)R_n}^{R_n} \hat{\rho}_n(y,k_n) e^{-\alpha y} \dd y
	&= \bigO{1} \hat{\rho}_n((1-\varepsilon^\ast)R_n,k_n) k_n^s e^{-\alpha(1-\varepsilon^\ast)R_n} \\
	&= \bigO{\hat{\rho}_n((1-\varepsilon^\ast)R_n,k_n) k_n^s n^{-2\alpha(1 - \varepsilon^\ast)}} \\
	&= \smallO{ \hat{\rho}_n((1-\varepsilon^\ast)R_n,k_n) n^{\frac{s}{2\alpha + 1} - 2\alpha(1-\varepsilon^\ast)}} = \smallO{1},
\end{align*}
where the last step follows by our choice of $\varepsilon^\ast$. Hence it is enough to prove both statement on $\Rcal_n((1-\varepsilon)R_n, R_n)$ instead of $\Rcal_n$.
\PvdH{@All: This is where we use that $s < 2\alpha(2\alpha + 1)$. I think this can be removed by using a more careful bound but I am not sure this is needed.}

\paragraph{Proof of statement 1.}
Since we already showed that
\[
	\lim_{n \to \infty} k_n^s \int_{(1-\varepsilon)R_n}^{R_n} \hat{\rho}_n(y,k_n) e^{(\beta-\alpha) y} \dd y = 0,
\]
by our assumption on $h_n(y)$, it is enough to show that for sufficiently large $C > 0$,
\begin{equation}\label{eq:concentration_argument_an_-}
	\lim_{n \to \infty} k_n^s \int_0^{a_n^-} \hat{\rho}_n(y,k_n) e^{(\beta-\alpha) y} \dd y = 0,
\end{equation}
and
\begin{equation}\label{eq:concentration_argument_an_+}
	\lim_{n \to \infty} k_n^s \int_{a_n^+}^{(1-\varepsilon)R_n} \hat{\rho}_n(y,k_n) e^{(\beta-\alpha) y} \dd y = 0.
\end{equation}
%\TM{ Why can we change the upper limit of the integral? PLease elaborate (in the paper) }

For simplicity we write $\mu(y) := \mu_{\alpha,\nu}\left(\BallPo{y}\right)$.. Now fix some $0 < \delta < 1$ and let $n$ be large enough such that 
\begin{enumerate}
\item $\sup_{0 < y \le (1-\varepsilon)R_n} |\phi_n(y)| < \delta$,
\item $(\ell_n + C \kappa_n)(1 - \delta) > k_n$ and
\item $(\ell_n - C \kappa_n)(1 + \delta) < k_n$
\end{enumerate}
